---
output:
  word_document: default
  html_document: default
---
# Mack Ford
## Model Validation - Answer
### 2/4/19

```{r}
#loading packages
#install.packages("caret")
library(caret)
library(MASS)
library(tidyverse)

#reading in Data
bike<- read_csv("hour.csv")

#updating data
bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4"))

#Updating Year to Factor
bike = bike %>% mutate(yr = as_factor(as.character(yr)))

#Updating mnth to factor
bike = bike %>% mutate(mnth = as_factor(as.character(mnth)))

#Updating hour to factor
bike = bike %>% mutate(hr = as_factor(as.character(hr)))

#updating Holiday to a factor and renaming
bike = bike %>% mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"NotHoliday" = "0",
"Holiday" = "1"))

#updating Working Day to a factor and renaming
bike = bike %>% mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1"))

#updating weathersit to a factor and renaming
bike = bike %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2",
"LightPrecip" = "3",
"HeavyPrecip" = "4"))

#updating weekday to a factor and renaming
bike = bike %>% mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" = "4",
"Friday" = "5",
"Saturday" = "6",
"Sunday" = "0"))

```

Now we are going to set up training and testing datasets

```{r}
#Task 1
#caret create data partition with a y variable of count. Setting the seed of 1234. Then splitting the sets
train.rows <- createDataPartition(y=bike$count, p=0.7, list = FALSE)
set.seed(1234)
train<- bike[train.rows,]
test <- bike[-train.rows,]
```

Task 2: 


- Train has 12167 Observations with 17 variables


- Test has 5212 observations with 17 variables

```{r}
#Task 3 - Linear Regression Model
Model1 <- lm(count~ season + mnth +hr + holiday + weekday + temp + weathersit, train)
summary(Model1)
```

This model has an adjusted R Squared of .62. This leads me to believe this is a good model. it also has a pvalue less than .05 so we can say these values are good predictors of Count

```{r}
#Task 4 - building predictions
trainingPredictions <- predict(Model1, newdata = train)
head(trainingPredictions)
```


The first 4 predictions are negative. This is note what I was expecting to see from the model, but then we go into the positives with 147 and 299. 

```{r}
#Task 6 - building predictions Test
testPredictions <- predict(Model1, newdata = test)
head(testPredictions)
```

The predictions are similar to the training dataset. This is leads me to believe that our model will handle data that it has not seen before, well. Next we will calculate th Rsquared

```{r}
#Task 6 Rsquared
SSE = sum((test$count - testPredictions)^2) #sO^2 from model
SST = sum((test$count - mean(test$count))^2) #sO^2 "naive" model
1 - SSE/SST 

```

Task 6 above gives us a R-Squared value of .619, call it .62. This is very similar to the Adjusted R-Squared value of the model with our training data. This means our model should behave in the "real world" when it sees new data. 

Task 7

The way I understand Kfolds. You pick a number of splits you want to have. typically they are 3, 5 or 10. Then you build out "K" number of models. Say you picked 5. You would then have 5 models with 5 different training sets (see below). This is different from the Test/Train Split. The test train split, you only build one model. The risk here is, you are building a model and are not using all of your data. 

###Example for KFold


####Model 1 


- Training Set 1: block 1, 2, 3, 4 


- TesT Set 1: block 5


####Model 2


- Training Set 2: block 1, 2, 3, 5 


- TesT Set 2: block 4


####Model 3


- Training Set 3: block 1, 2, 4, 5 


- TesT Set 3: block 3


####Model 4


- Training Set 4: block 1, 3, 4, 5 


- TesT Set 4: block 2


####Model 5


- Training Set 5: block 2, 3, 4, 5 


- TesT Set 5: block 1


###Example for Train/Test Split


- Training Set: 80% of your data


- Test Set: 20% of your data




