---
output:
  word_document: default
  html_document: default
---
#Mack Ford
##Assignment One - Simple Linear Regression & Correlation
###1/28/2019



```{r}
library(tidyverse)
#install.packages("GGally")
library(GGally)

#Task 1
air <-airquality
glimpse(air)
summary(air)
```

a. This data describes air quality. It contains month, day, Temp, wind, Solar R, and Ozone 

b. This data set contains 153 observations and 6 Variables

c. Yes there are missing variables, in Ozone and Solar R

d. Ozone is most likely to be the response variable. 

```{r}
#Task 2
#deleting the data 
air2 = air %>% filter(!is.na(Ozone)) %>% filter(!is.na(Solar.R))
```

Now that we have deleted the NA we only have 111 observations. 

```{r}
# Task 3
ggpairs(air2)

ggcorr(air2, label = TRUE)
```

a. Temperature is the strognest correlated variable with Ozone
b. Day has the least strongest correlation with Ozone

```{r}
#Task 4 - Plotting temp and Ozone

ggplot(air2, aes(Temp, Ozone)) +
  geom_point()+labs(title ="Temp Vs. Ozone") + 
  geom_smooth(method = "lm", se = FALSE, color = "red")+theme_bw()
```

Temperature vs. Ozone - The data has a positive correlation (.7). We can see that in the scatterplot that there is an upward trend.

```{r}
#Task 5 Linear Model
model1<- lm(Ozone~Temp,air2)
#Summary tells you the P value, intercepts, and R^2
summary(model1)


```

R^2 Value of .48, and a Pvalue or less than .05. The Pvalue lets us know that Temp is a significant predictor of Ozone. The R squared value of .48 shows that it is a decent model, but it could be improved. 

```{r}
#Task 5 Cont. 
confint(model1)
```

With a 95% confidence we can say the slop will land between 1.96 and 2.91

```{r}
#Task6 - Repeat 4
ggplot(air2, aes(Temp, Ozone)) +
  geom_point()+labs(title ="Temp Vs. Ozone") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +theme_bw()
```

```{r}
#Task 7 
#doing the Predicting
Testdata<-data.frame(Temp =c(80))
predict(model1, newdata = Testdata, interval = "predict")
#Prediction Intervals to the original Data Set
temp_var<-predict(model1, interval = "prediction")
new_df<-cbind(air2, temp_var)
```

we are predicting that with a temperatrue of 80 we will have an Ozone of 47.48. lower and Upper prediction iterval for the temperature is: lower = -.15 & Upper = 95.11


```{r}
#Task 8
ggplot(air2, aes(Day, Ozone))+
  geom_point() + labs(title = "Day vs. Ozone") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + theme_bw()
```

The relationship between Day and Ozone does not seem to be correlated. Our trend line is almost 0.

```{r}
#Task 9
model2<- lm(Ozone~Day, air2)
summary(model2)
```

based off the summary of model 2. We have a poor RSquared value, but we do have a p-value that is less that .05 which tells us we have a significant predictor of Ozone. 

```{r}
#Task 9 Cont
confint(model2)
```

The slope will typically fall between -.74 and .70


```{r}
#Task 10 
ggplot(air2, aes(Day, Ozone))+
  geom_point() + labs(title = "Day vs. Ozone") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + theme_bw()
```

